---
title: "Introduction to metacor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to metacor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(metacor)
```

# What `metacor` is for

Meta-analyses of pre--post studies require summary statistics that are
rarely reported in primary papers --- in particular the **standard
deviation of the change scores** (`sd_change`) and the **pre--post
correlation coefficient** (`r`). Without these, standardised mean
differences (SMDs) and their variances cannot be estimated correctly,
and the usual fallback is a sensitivity analysis that not every team
performs.

`metacor` automates the calculation and imputation of `sd_change` and
`r` from whatever summary statistics the primary studies *did* report
(p-values, confidence intervals, sample sizes), and produces a
transparent imputation report so the chosen strategy is fully
documented.

The package is designed for two designs:

- **Single-group pre--post** (one arm, treatment only): set
  `single_group = TRUE`.
- **Pre--post with control** (intervention vs. control): the default.

# Input data: the canonical schema

`metacor_dual()` expects a data frame with these columns (suffix
`_int` for the intervention group, `_ctrl` for the control group):

| Column            | Meaning                                                |
|-------------------|--------------------------------------------------------|
| `study_name`      | identifier for each row                                |
| `mean_pre_*`      | pre-test mean                                          |
| `mean_post_*`     | post-test mean                                         |
| `sd_pre_*`        | pre-test SD                                            |
| `sd_post_*`       | post-test SD                                           |
| `n_*`             | sample size                                            |
| `p_value_*`       | within-group pre--post p-value (optional)              |
| `upper_ci_*`      | upper bound of the 95% CI for the change (optional)    |
| `lower_ci_*`      | lower bound of the 95% CI for the change (optional)    |

In single-group designs only the `_int` columns are needed.

## Flexible column matching

You don't have to use the exact canonical names. By default
`metacor_dual()` accepts:

- **Case and separator variants.** `MeanPre_Int`, `mean.pre.int`,
  `meanpreint` are all silently treated as `mean_pre_int`.
- **A closed dictionary of synonyms.** `baseline_mean_treatment`
  resolves to `mean_pre_int`; `_placebo` and `_sham` resolve to `_ctrl`;
  `pval` resolves to `p_value`; `m1` and `m2` resolve to `mean_pre`
  and `mean_post`.

Every flexible match is reported. You can disable this with
`column_matching = "strict"` if you want the function to require canonical
names exactly.

# A minimal single-group example

```{r}
df <- data.frame(
  study_name    = paste0("Study", 1:9),
  p_value_int   = c(1.04e-07, NA, NA, NA, NA, 0.021, NA, NA, NA),
  n_int         = c(10, 10, 10, 10, 15, 15, 10, 10, 10),
  mean_pre_int  = c(8.17, 10.09, 10.18, 9.85, 9.51, 7.70, 10.00, 11.53, 11.20),
  mean_post_int = c(10.12, 12.50, 12.56, 10.41, 10.88, 9.20, 10.80, 13.42, 12.00),
  sd_pre_int    = c(1.83, 0.67, 0.66, 0.90, 0.62, 0.90, 0.70, 0.60, 1.90),
  sd_post_int   = c(1.85, 0.72, 0.97, 0.67, 0.76, 1.10, 0.70, 0.80, 1.80),
  upper_ci_int  = NA_real_,
  lower_ci_int  = NA_real_
)
```

Run `metacor_dual()` with `effect_size = "smd_pre"` (the default) and
the `cv` imputation method:

```{r}
result <- metacor_dual(
  df,
  single_group = TRUE,
  effect_size  = "smd_pre",
  impute_method = "cv",
  digits       = 3,
  verbose      = FALSE
) |> suppressWarnings()
result[, c("study_name", "r_int", "sd_change_int",
           "smd_pre_int", "var_smd_pre_int", "se_smd_pre_int")]
```

The function added six columns to your data: the imputed pre--post
correlation `r_int`, the SD of the change `sd_change_int`, the SMD
itself `smd_pre_int`, and its variance / standard error.

# Choosing an effect size

`effect_size` selects one of four families:

| Value               | What it computes                                                   | Denominator   |
|---------------------|---------------------------------------------------------------------|---------------|
| `"smd_pre"` (default)| SMD using pre-test SD as denominator                              | `sd_pre_int`  |
| `"smd_change"`       | SMD using SD of change scores                                      | `sd_change_int`|
| `"smd_pooled"`       | Standardised change mean difference, pooled SD (controlled designs)| pooled SD     |
| `"smd_diff_groups"`  | Difference of standardised changes between intervention and control | per-group SD  |

The variance formulas differ accordingly. `smd_pre` requires `r`;
`smd_change` does not.

# Imputation methods for `sd_change`

When `sd_change` cannot be computed from the reported p-value or CI,
`impute_method` chooses the imputation strategy:

| Method     | What it does                                                                          |
|------------|----------------------------------------------------------------------------------------|
| `"none"`   | leaves missing `sd_change` (and downstream `r`) as NA                                  |
| `"direct"` | replaces missing values with the **maximum** observed `sd_change` (conservative)       |
| `"mean"`   | replaces missing values with the **mean** of observed `sd_change`                      |
| `"cv"`     | uses `\|mean_diff\|` × **robust global median CV** (recommended; see Higgins et al. 2023) |

Manual imputation per study is also supported via
`custom_sd_change_int` and `custom_sd_change_ctrl` (lists of
`list(row = ..., value = ...)`).

# Imputation report

Set `report_imputations = TRUE` to generate a Word file
(`imputation_report.docx`) describing every imputation performed: which
study, which group, the imputed value, the suggested feasible range, and
the method used. This file is meant to be included as supplementary
material in your meta-analysis.

```{r, eval = FALSE}
result <- metacor_dual(df, single_group = TRUE, impute_method = "cv",
                       report_imputations = TRUE)
# Writes "imputation_report.docx" in the working directory.
```

# Internal consistency checks

After running `metacor_dual()`, you can pass the result to
`check_metacor_consistency()` to flag suspicious data points:

```{r, eval = FALSE}
result_with_md <- metacor_dual(df, single_group = TRUE,
                               impute_method = "cv",
                               mean_differences = TRUE,
                               verbose = FALSE)
result_checked <- check_metacor_consistency(result_with_md, interpret = TRUE)
result_checked$summary_int
```

The function adds logical flag columns (`flag_p_mismatch_int`,
`flag_CI_mismatch_int`, `flag_sd_change_out_of_range_int`,
`flag_r_extreme_int`) and, with `interpret = TRUE`, narrative
explanations per study.

# Migrating from `metacor` 1.2.x

If you are upgrading from a previous version, see the companion
vignette `vignette("migrating-from-1.2", package = "metacor")` for the
mapping of old argument names, old `effect_size` values
(`"SMDpre"`, `"ScMDpooled"`, ...), and old column names (`sd_diff_*`,
`_Con`, ...) to their v1.3 canonical equivalents. All v1.2.x names
still work, with deprecation warnings.

# References

Higgins, J. P. T., Thomas, J., Chandler, J., Cumpston, M., Li, T., Page,
M. J., & Welch, V. A. (Eds.). (2023). *Cochrane handbook for systematic
reviews of interventions* (Version 6.3). Cochrane.
<https://www.cochrane.org/authors/handbooks-and-manuals/handbook/current>

Fu, R., Vandermeer, B. W., Shamliyan, T. A., O'Neil, M. E., Yazdi, F.,
Fox, S. H., & Morton, S. C. (2013). *Handling Continuous Outcomes in
Quantitative Synthesis*. Methods Guide for Comparative Effectiveness
Reviews. AHRQ Publication No. 13-EHC103-EF.
<https://effectivehealthcare.ahrq.gov/reports/final.cfm>

# Session info

```{r}
sessionInfo()
```
