---
title: "Migrating from metacor 1.2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Migrating from metacor 1.2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(metacor)
```

# Why this rename happened

`metacor` 1.2.x accumulated three naming inconsistencies that made the
package harder to use and to write about:

1. **`SDdiff` vs `SDchange`.** The same concept --- the standard deviation
   of pre--post change scores --- was called `sd_diff_int` in code and
   `SDchange` in the manuscript. The Cochrane Handbook canonical term is
   the latter, so version 1.3 standardises on **`sd_change`**
   throughout.
2. **Mixed casing.** `MeanDifferences` (CamelCase), `SMD_method`
   (PascalCase), `sd_pre_Int` (mixed) and `apply_hedges` (snake_case)
   coexisted. Version 1.3 uses **pure snake_case** for everything.
3. **Internal mismatch.** The effect size `SMDchange` was computed using
   a denominator named `sd_diff_int`. After the rename, `smd_change` is
   computed from `sd_change_int` --- one root word per concept.

This vignette tells you how to migrate. **Your existing 1.2.x code keeps
working.** Every old argument and every old column name is recognised by
the column matcher and translated to the new schema, with a
`lifecycle::deprecate_warn()` so the path forward is visible.

# At a glance

```{r, eval = FALSE}
# v1.2.x
result <- metacor_dual(
  df,
  method               = "both",
  SMD_method           = "SMDpre",
  MeanDifferences      = TRUE,
  custom_sd_diff_int   = list(list(row = 7, value = 0.5)),
  single_group         = TRUE
)

# v1.3.0 (equivalent)
result <- metacor_dual(
  df,
  derive_from           = "both",
  effect_size           = "smd_pre",
  mean_differences      = TRUE,
  custom_sd_change_int  = list(list(row = 7, value = 0.5)),
  single_group          = TRUE
)
```

The two calls produce **bit-for-bit identical numerical output** (only
column names differ). This was verified across 23 golden test fixtures
covering every effect-size × imputation combination.

# Argument renames

| v1.2.x argument        | v1.3.0 argument         | Notes                                  |
|------------------------|--------------------------|----------------------------------------|
| `method`               | `derive_from`            | Three args contained "method" before. |
| `SMD_method`           | `effect_size`            | Describes what the user picks.        |
| `MeanDifferences`      | `mean_differences`       | snake_case throughout.                |
| `custom_sd_diff_int`   | `custom_sd_change_int`   | "diff" → "change".                    |
| `custom_sd_diff_con`   | `custom_sd_change_ctrl`  | Also "con" → "ctrl".                  |
| `apply_hedges`, `impute_method`, `single_group`, `report_imputations`, `verbose`, `digits`, `add_to_df` | unchanged | |
| *(new)*                | `column_matching`        | `"flexible"` (default) or `"strict"`. |

# Effect-size value renames

The values accepted by `effect_size` (formerly `SMD_method`) also moved
to snake_case:

| v1.2.x value    | v1.3.0 value         |
|-----------------|----------------------|
| `"SMDpre"`      | `"smd_pre"`          |
| `"SMDchange"`   | `"smd_change"`       |
| `"ScMDpooled"`  | `"smd_pooled"`       |
| `"ScMDpre"`     | `"smd_diff_groups"`  |

The new names are more descriptive: `smd_pooled` is "SMD with pooled
SD", `smd_diff_groups` is "difference of SMDs between groups".

# Column-name renames

## Input columns

| v1.2.x        | v1.3.0          |
|---------------|-----------------|
| `meanPre_Int` | `mean_pre_int`  |
| `meanPost_Int`| `mean_post_int` |
| `sd_pre_Int`  | `sd_pre_int`    |
| `sd_post_Int` | `sd_post_int`   |
| `n_Int`       | `n_int`         |
| `p_value_Int` | `p_value_int`   |
| `upperCI_Int` | `upper_ci_int`  |
| `lowerCI_Int` | `lower_ci_int`  |
| same with `_Con` | same with `_ctrl` |

You **don't have to rename your columns manually**. The flexible column
matcher (active by default) accepts:

- The new canonical names (`mean_pre_int`, ...).
- The v1.2.x mixed-case names (`meanPre_Int`, ...) --- silent rename via
  case normalisation.
- `_Con` is treated as a deprecated alias for `_ctrl` and emits a
  deprecation warning.
- A closed dictionary of synonyms: `_intervention`, `_treatment`, `_trt`,
  `_exp`, `_placebo`, `_pbo`, `_sham`, `_passive`, `m1`/`mean1`,
  `m2`/`mean2`, `pval`, `ci_upper/lower`, etc. See the package
  documentation for the full list.

## Output columns

These are the columns `metacor_dual()` *adds* to your data:

| v1.2.x output            | v1.3.0 output           |
|--------------------------|--------------------------|
| `r_int`                  | `r_int` *(unchanged)*    |
| `r_con`                  | `r_ctrl`                 |
| `sd_diff_int`            | `sd_change_int`          |
| `sd_diff_con`            | `sd_change_ctrl`         |
| `pct_change_int`         | `pct_change_int` *(unchanged)* |
| `pct_change_con`         | `pct_change_ctrl`        |
| `SMDpre_int`             | `smd_pre_int`            |
| `varSMDpre_int`          | `var_smd_pre_int`        |
| `seSMDpre_int`           | `se_smd_pre_int`         |
| `SMDchange_*`            | `smd_change_*`           |
| `ScMDpooled`             | `smd_pooled`             |
| `varScMDpooled`          | `var_smd_pooled`         |
| `seScMDpooled`           | `se_smd_pooled`          |
| `ScMDpre`                | `smd_diff_groups`        |
| `varScMDpre`             | `var_smd_diff_groups`    |
| `seScMDpre`              | `se_smd_diff_groups`     |
| `meanDiff_int`           | `mean_diff_int`          |
| `varMeanDiff_int`        | `var_mean_diff_int`      |
| `seMeanDiff_int`         | `se_mean_diff_int`       |

For `check_metacor_consistency()`, the flag column
`flag_sd_diff_out_of_range_*` becomes `flag_sd_change_out_of_range_*`,
and `summary_con` becomes `summary_ctrl`.

# Strict mode

If you want to lock down your code against any flexible matching --- for
example to ensure a published pipeline never silently absorbs a typo ---
pass `column_matching = "strict"`:

```{r, eval = FALSE}
result <- metacor_dual(df, column_matching = "strict")
```

In strict mode, only the exact canonical column names are accepted; any
deviation passes through untouched and the function will error if a
required column is missing.

# Removal timeline

The compatibility layer is provisional. The plan is:

- **1.3.x series**: legacy names continue to work with deprecation
  warnings. Update at your own pace.
- **2.0.0**: legacy names are removed. Code that ignored the
  deprecation warnings will break.

There is no firm date for 2.0.0 yet --- it will land at least 6 months
after 1.3.0 reaches CRAN to give users time to update.

# A side-by-side example

```{r, eval = FALSE}
df <- data.frame(
  study_name   = c("S1", "S2", "S3"),
  meanPre_Int  = c(8.0, 10.0, 11.0),    # v1.2.x mixed case
  meanPost_Int = c(10.0, 12.0, 13.0),
  sd_pre_Int   = c(1.5, 0.7, 0.9),
  sd_post_Int  = c(1.6, 0.8, 1.0),
  n_Int        = c(12, 12, 12),
  p_value_Int  = c(0.001, NA, 0.01),
  upperCI_Int  = NA_real_,
  lowerCI_Int  = NA_real_
)

# This call (v1.2.x style) still works in 1.3.0:
res_legacy <- metacor_dual(df,
                           SMD_method      = "SMDpre",
                           method          = "both",
                           MeanDifferences = TRUE,
                           single_group    = TRUE)
#> Warning: `metacor_dual(method)` was deprecated in metacor 1.3.0.
#> Please use `derive_from` instead.
#> Warning: `metacor_dual(SMD_method)` was deprecated ...
#> ...

# The recommended v1.3.0 form, using canonical column names:
df2 <- df
names(df2) <- c("study_name", "mean_pre_int", "mean_post_int",
                "sd_pre_int", "sd_post_int", "n_int", "p_value_int",
                "upper_ci_int", "lower_ci_int")
res_canonical <- metacor_dual(df2,
                              effect_size      = "smd_pre",
                              derive_from      = "both",
                              mean_differences = TRUE,
                              single_group     = TRUE)

identical(res_legacy[setdiff(names(res_canonical), names(res_legacy))],
          res_canonical[setdiff(names(res_canonical), names(res_legacy))])
#> The output columns are renamed; the numerical values are identical.
```

# Session info

```{r}
sessionInfo()
```
